---
title: "First Draft"
author: "Jack de la Parra"
date: "6/23/2020"
output: html_document
---
Loading the initial dataset:
```{project setup, include=TRUE}
ncaa_volleyball <- read.csv("http://www.stat.cmu.edu/cmsac/sure/materials/data/regression_projects/womens_ncaa_volleyball_20182019.csv")
```


Now attempting to do an EDA, where I will first begin to look at how certain variables correlate with wins
```{r}
ncaa_volleyball %>%
  arrange(desc(w))

install.packages("GGally")
library(GGally)
library(tidyverse)

aces <- ncaa_volleyball %>%
  ggplot(aes(x = aces_per_set, y = w)) +
  geom_point() +
  theme_bw()
plot(aces)

cor(x = ncaa_volleyball$aces_per_set, y = ncaa_volleyball$w)
ggpairs(ncaa_volleyball, columns = c("aces_per_set", "w"))

# Exploring correlation coefficients as well as histograms of variables to see normality
# somewhat strong correlation at around  r = .330


ncaa_volleyball %>%
  ggplot(aes(x = assists_per_set, y = w)) +
  geom_point() +
  theme_bw()

ggpairs(ncaa_volleyball, columns = c("assists_per_set", "w"))

# very high correlation, around .780


ncaa_volleyball %>%
  ggplot(aes(x = blocks_per_set, y = w)) +
  geom_point() +
  theme_bw()
ggpairs(ncaa_volleyball, columns = c("blocks_per_set", "w"))

# decently high correlation of around .492

ncaa_volleyball %>%
  ggplot(aes(x = digs_per_set, y = w)) +
  geom_point() +
  theme_bw()
ggpairs(ncaa_volleyball, columns = c("digs_per_set", "w"))

# comparatively low correlation, around .254


# Group decided using win percentage instead of wins might be more effective for comparisons, since each team played a different number of games

ncaa_volleyball %>%
  ggplot(aes(x = aces_per_set, y = win_pct)) +
  geom_point() +
  theme_bw()
ggpairs(ncaa_volleyball, columns = c("aces_per_set", "win_pct"))
# slightly higher correlation now, around .339 rather than ~.330

ncaa_volleyball %>%
  ggplot(aes(x = assists_per_set, y = win_pct)) +
  geom_point() +
  theme_bw()
ggpairs(ncaa_volleyball, columns = c("assists_per_set", "win_pct"))
# correlation now even higher, around .798

ncaa_volleyball %>%
  ggplot(aes(x = blocks_per_set, y = win_pct)) +
  geom_point() +
  theme_bw()
ggpairs(ncaa_volleyball, columns = c("blocks_per_set", "win_pct"))
# now slightly higher correlation around .507

ncaa_volleyball %>%
  ggplot(aes(x = digs_per_set, y = win_pct)) +
  geom_point() +
  theme_bw()
ggpairs(ncaa_volleyball, columns = c("digs_per_set", "win_pct"))
# slightly higher, but relatively low correlation of around .261

ncaa_volleyball %>%
  ggplot(aes(x = kills_per_set, y = win_pct)) +
  geom_point() +
  theme_bw()
ggpairs(ncaa_volleyball, columns = c("kills_per_set", "win_pct"))
# very high correlation here of 0.824

ncaa_volleyball %>%
  ggplot(aes(x = hit_pct, y = win_pct)) +
  geom_point() +
  theme_bw()
ggpairs(ncaa_volleyball, columns = c("hit_pct", "win_pct"))
# highest correlation so far, of around 0.852

# now I will mutate a new variable, attacks per set, in attempt to see if more aggressive offensive teams are more successful
library(tidyverse)
ncaa_volleyball_attacks <- ncaa_volleyball %>%
  mutate(attacks_per_set = total_attacks / s)

cor(x = ncaa_volleyball_attacks$attacks_per_set, y = ncaa_volleyball_attacks$win_pct)
# There is very low correlation, which implies that attacks and wins are not very strongly related

library(GGally)
ggpairs(ncaa_volleyball_attacks, columns = c("digs_per_set", "attacks_per_set"))
```

Beginning to create some potential models to fit the data set in attempts to predict win percentage

```{r}
init_lm <- lm(win_pct ~ assists_per_set,  ncaa_volleyball)

init_lm$coefficients

summary(init_lm)

# Both p-values are indeed significant, the standard errors appear to be low as well, but perhaps the predictions do not matchup very well with the actual data
```
```{r}
ncaa_volleyball_init <- ncaa_volleyball %>%
  mutate(init_preds = predict(init_lm)) 

ncaa_volleyball_init %>%
  ggplot(aes(x = win_pct, y = init_preds)) +
  geom_point(alpha = 0.5) +
  theme_bw() +
  labs(x = "Actual Win Percentage", 
       y = "Predicted Win Percentage")
# This appears to be pretty highly correlated and upon running the following function, we can observe a correlation of 0.7978 which is very strong.
cor(x = ncaa_volleyball_init$win_pct, y = ncaa_volleyball_init$init_preds)
```
Now I attempt to make a multi-variable model, utilizing variables from different facets of the game, defense, offense, and secondary offense
```{r}
multi_lm <- lm(win_pct ~ assists_per_set + blocks_per_set + kills_per_set, ncaa_volleyball)

summary(multi_lm)
# They have low p-values and relatively low standard errors. This has a higher adjusted R-squared value as well (which accounts for number of variables included in the model)
```
```{r}
ncaa_volleyball_multi <- ncaa_volleyball %>%
  mutate(multi_preds = predict(multi_lm)) 

ncaa_volleyball_multi %>%
  ggplot(aes(x = win_pct, y = multi_preds)) +
  geom_point(alpha = 0.5) +
  theme_bw() +
  labs(x = "Actual Win Percentage", 
       y = "Predicted Win Percentage")



# This appears to be very strongly correlated, and we can observe a correlation of 0.8637 by running the following function

cor(x = ncaa_volleyball_multi$win_pct, y = ncaa_volleyball_multi$multi_preds)
```



Note: I had originally begun this project in an RScript and then had to move it over to R Markdown file, creating some flaws in formatting